---
title: "Untitled"
author: "Paulo"
date: "02-07-2015"
output: html_document
---

Respostas dos exercícios de "Advanced R". Utilizar apenas como guia de estudo.

### 2.1.3 Exercises

1. What are the six types of atomic vector? How does a list differ from an
   atomic vector?

    Six types of atomic vector: character, integer, double, logical, complex and raw.
    Atomic vectors permit elements of only one type. Lists can have elements of any type together, including lists.

1. What makes `is.vector()` and `is.numeric()` fundamentally different to
   `is.list()` and `is.character()`?
    
    `is.vector()` does not test if object is of type vector. It returns `TRUE` when the argument is a vector without attributes (excluding names).
    
1. Test your knowledge of vector coercion rules by predicting the output of
   the following uses of `c()`:

    ```{r, eval=FALSE}
    c(1, FALSE) 
    ```
    
      The first element is of type integer, more flexible than logical. So the   second element will be coerced to type integer - `0`.


    ```{r, eval=FALSE}
    c("a", 1)  
    ```
    The more flexible here is character type, so the second element will be coerced to `"1"`.

    ```{r, eval=FALSE}
    c(list(1), "a")
    ```
    This will coerce `"a"` to a list and combine both elements into one list.

    ```{r, eval=FALSE}
    c(TRUE, 1L)
    ```
    Integer is the most flexible type. `TRUE` will be coerced to `1`.
    
1.  Why do you need to use `unlist()` to convert a list to an 
    atomic vector? Why doesn't `as.vector()` work? 

    `as.vector()` and does not try to coerce objects to vector, but coerces objects to the same type, removing attributes if objecs are atomic.

1. Why is `1 == "1"` true? Why is `-1 < FALSE` true? Why is `"one" < 2` false?
    According to the manual: "If the two arguments are atomic vectors of different types, one is coerced to the type of the other, the (decreasing) order of precedence being character, complex, numeric, integer, logical and raw."
    
    In `1 == "1"`, `1` is coerced to `"1"` (character), and the expression returns `TRUE`. In `-1 < FALSE`, `FALSE` is coerced to `0` (integer), and thus the expression is true. In the last case, `"one" < 2`, `2` is coerced to `"2"` (character), and the expression is false
    

1. Why is the default missing value, `NA`, a logical vector? What's special about logical vectors? (Hint: think about `c(FALSE, NA_character_)`.)
    Logical vectors are the less flexible type of vector. Thus, NA is a logical vector so that it can be coerced to the correct type (`NA_real_`, `NA_integer_` or `NA_character_`) when in a vector with elements of other type.
    
# 2.2.2 Exercises


1.  An early draft used this code to illustrate `structure()`:

    ```{r}
    structure(1:5, comment = "my attribute")
    ```

    But when you print that object you don't see the comment attribute.
    Why? Is the attribute missing, or is there something else special about
    it? (Hint: try using help.)
    
    **Answer** `?attributes` show that `comment` is a special attribute, and have restrictions on values to be set. 
    
1.  What happens to a factor when you modify its levels? 
    
    ```{r, results = "none"}
    f1 <- factor(letters)
    levels(f1) <- rev(levels(f1))
    ```

    **Answer** The factor gets modified according to the levels modification.

1.  What does this code do? How do `f2` and `f3` differ from `f1`?

    ```{r, results = "none"}
    f2 <- rev(factor(letters))

    f3 <- factor(letters, levels = rev(letters))
    ```
    
    **Answer** `f1` is a factor vector composed of letters and, as it's levels are reversed, the vector is also modified.
    `f2` is a factor vector created by reversing a vector just like the original f1. In this case, levels are in crescent order, since they weren't modified.
    `f3` is also a factor vector containing `letters` in the crescent order, but with the levels in decreasing order.
    
# 2.3.1 Exercises

1.  What does `dim()` return when applied to a vector?

    **Answer** `NULL`. `dim()` is an attribute to multi-dimensional vectors (matrices and arrays).

1.  If `is.matrix(x)` is `TRUE`, what will `is.array(x)` return?
    **Answer** `TRUE`. Matrices are a special case of arrays, have only two dimensions. Probably for being most commonly used, matrices have its own name.

1.  How would you describe the following three objects? What makes them
    different to `1:5`?

    ```{r}
    x1 <- array(1:5, c(1, 1, 5))
    x2 <- array(1:5, c(1, 5, 1))
    x3 <- array(1:5, c(5, 1, 1))
    ```
    **Answer** Object `x1` could be called an array vector, with all elements along the third dimension.  `x2` would be an array equivalent to a row vector and `x3` could be compared to a column vector. They differ from a plain vector (`1:5`) in the `dim` attribute and consequently in the number of dimensions (3).
    
# 2.4.5 Exercises

1.  What attributes does a data frame possess?
    
    **Answer** It has properties of matrices and lists: `names()`, `colnames()`, `rownames()`, `length()`, `ncol`, `nrow`, `dim` and `class`
    
1.  What does `as.matrix()` do when applied to a data frame with 
    columns of different types?

    **Answer** It tries coercing all columns to the most flexible type in the data frame.

1.  Can you have a data frame with 0 rows? What about 0 columns?

    **Answer** It is possible to have an empty data frame with both 0 columns *and* 0 rows by executing only `data.frame()`. However it is impossible to have these characteristics separately.